### 2010-07-12 papers/icfp-2010-desc/paper_desc_levitation.tex
Desc@ICFP: fix typo

Note that up to isomorphism, $\SYMBDIndx$ is subsumed by
$\DHindx{\Unit}{\!}$. However, the apparent duplication has some
value.  Unlike its counterpart, $\SYMBDIndx$ is first-order: we prefer
not to demand dummy functions from \(\Unit\) in ordinary data,
e.g. \(\NatSuc{(\LAM{\_}n)}\). It is na{\"\i}ve to imagine that up to
isomorphism, any representation of data will do.  First-order
representations are finitary by construction, and thus admit a richer,
componentwise decidable equality than functions may in
general possess.\footnote{E.g., extensionally, there is one inhabitant of
\(\EnumT{\void}\To\Nat\); intensionally, there is a countable infinitude
which it is not safe to collapse.}

>>>>

Note that up to isomorphism, $\SYMBDIndx$ is subsumed by
$\DHindx{\Unit}{\!}$. However, the apparent duplication has some
value.  Unlike its counterpart, $\SYMBDIndx$ is first-order: we prefer
not to demand dummy functions from \(\Unit\) in ordinary data,
e.g. \(\NatSuc{(\LAM{\_}n)}\). It is na{\"\i}ve to imagine that up to
isomorphism, any representation of data will do.  First-order
representations are finitary by construction, and thus admit a richer,
componentwise decidable equality that functions may in
general possess.\footnote{E.g., extensionally, there is one inhabitant of
\(\EnumT{\void}\To\Nat\); intensionally, there is a countable infinitude
which it is not safe to collapse.}

### 2010-07-12 papers/icfp-2010-desc/paper_idesc.tex
Desc@ICFP: fix typo

However, so far, we are only able to define and manipulate
\emph{closed} terms. By abstracting over $\SYMBVal$, it is possible to
build and manipulate \emph{open} terms, that is, terms in a
context. The context is represented by a snoc-list of types:
%
\[
\stk{
\data \Ctxt : \Set \where \\
\;\;\begin{array}{@{}l@{\::\:\:}l@{\quad}l}
    \SYMBCtxtEmpty  & \Ctxt   \\
    \SYMBCtxtSnoc   & \Ctxt \To \Ty \To \Ctxt
\end{array}
}\]
%
An environment realises the context, associating a value for each type:
%
\[
\stk{
\SYMBEnv : \Ctxt \To \Set \\
\begin{array}{@{}l@{\:\:=\:\:}l}
\Env{\CtxtEmpty}        & \Unit \\
\Env{(\CtxtSnoc{G}{S})} & \TIMES{\Env{G}}{\Val{S}}
\end{array}
}\]
%
In this setting, we define $\SYMBVar$ by:
%
\[\stk{
\Var{}{} : \Ctxt \To \Ty \To \Set \\
\begin{array}{@{}ll@{\:\:=\:\:}l}
\Var{\CtxtEmpty}{& T}            & \Void \\
\Var{(\CtxtSnoc{G}{S})}{& T}     & \SUM{(\Var{G}{T})}{(S \Eq T)}
\end{array}
}\]
%
While $\SYMBVal$ was mapping the type to the corresponding host type,
$\SYMBVar$ indexes a value in the context, obtaining a proof that the
types matches. The $\SYMBlookup$ function precisely follow this
semantics:
%
\[\stk{
\SYMBlookup : \PI{\V{G}}{\Ctxt} 
          \Env{\V{G}} \To 
          \PI{\V{T}}{\Ty} 
          \Var{\V{G}}{\V{T}} \To
          \Val{\V{T}} \\
\begin{array}{@{}lll}
\lookup{\CtxtEmpty}{               \_}{                   \V{T}}{()}                    &   & \\
\lookup{(\CtxtSnoc{\V{G}}{.T})}{   \pair{\V{g}}{\V{t}}{}}{\V{T}}{(\SumRight{\C{refl}})} & = & \V{t} \\
\lookup{(\CtxtSnoc{\V{G}}{\V{S}})}{\pair{\V{g}}{\V{t}}{}}{\V{T}}{(\SumLeft{\V{x}})}     & = & \lookup{\V{G}}{\V{g}}{\V{T}}{\V{x}} 
\end{array}
}\]

>>>>

However, so far, we are only able to define and manipulate
\emph{closed} terms. By abstracting over $\SYMBVal$, it is possible to
build and manipulate \emph{open} terms, that is, terms in a
context. The context is represented by a snoc-list of types:
%
\[
\stk{
\data \Ctxt : \Set \where \\
\;\;\begin{array}{@{}l@{\::\:\:}l@{\quad}l}
    \SYMBCtxtEmpty  & \Ctxt   \\
    \SYMBCtxtSnoc   & \Ctxt \To \Ty \To \Ctxt
\end{array}
}\]
%
An environment realises the context, associating a value for each type:
%
\[
\stk{
\SYMBEnv : \Ctxt \To \Set \\
\begin{array}{@{}l@{\:\:=\:\:}l}
\Env{\CtxtEmpty}        & \Unit \\
\Env{(\CtxtSnoc{G}{S})} & \TIMES{\Env{G}}{\Val{S}}
\end{array}
}\]
%
In this setting, we define $\SYMBVar$ by:
%
\[\stk{
\Var{}{} : \Ctxt \To \Ty \To \Set \\
\begin{array}{@{}ll@{\:\:=\:\:}l}
\Var{\CtxtEmpty}{& T}            & \Void \\
\Var{(\CtxtSnoc{G}{S})}{& T}     & \SUM{(\Var{G}{T})}{(S \Eq T)}
\end{array}
}\]
%
While $\SYMBVal$ was mapping the type to the corresponding host type,
$\SYMBVar$ indexes a value in the context, obtaining a proof that the
types match. The $\SYMBlookup$ function precisely follow this
semantics:
%
\[\stk{
\SYMBlookup : \PI{\V{G}}{\Ctxt} 
          \Env{\V{G}} \To 
          \PI{\V{T}}{\Ty} 
          \Var{\V{G}}{\V{T}} \To
          \Val{\V{T}} \\
\begin{array}{@{}lll}
\lookup{\CtxtEmpty}{               \_}{                   \V{T}}{()}                    &   & \\
\lookup{(\CtxtSnoc{\V{G}}{.T})}{   \pair{\V{g}}{\V{t}}{}}{\V{T}}{(\SumRight{\C{refl}})} & = & \V{t} \\
\lookup{(\CtxtSnoc{\V{G}}{\V{S}})}{\pair{\V{g}}{\V{t}}{}}{\V{T}}{(\SumLeft{\V{x}})}     & = & \lookup{\V{G}}{\V{g}}{\V{T}}{\V{x}} 
\end{array}
}\]
